Correlation between 16S and shotgun

### 2019-11-08 scripts/20_correlation_phylum_bw_16S_shotgun.Rmd
log transform the relab

Two <- two %>% 
  pluck('shotgun') %>% 
  rename(shotgun_relab = relab) %>% 
  select(-Group) %>% 
    full_join(two %>% 
    pluck('s16S') %>% 
    rename(s16S_relab = relab) %>% 
    select(-Group)) %>%  
  # to uniform the two data
  mutate(s16S_relab = s16S_relab*100)

@@

```{r}
Two %>% 
  ggplot(aes(x = shotgun_relab, y = s16S_relab, col = phylum)) +
    geom_point() +
    facet_wrap(~sampleid) +
  scale_color_manual(values = brewer.pal(n = length(ol_phylum), name = "Dark2")) +
  labs(x = 'Shotgun data',
       y = '16S data',
       title = 'Relative abundance comparison between shotgun and 16S data\nat phylum level') +
  ggsave('../figs/correlation/cor_between_16_and_shotgun.jpg', width = 10, height = 10, dpi = 300)

@@

spearman_cor %>% 
  kable(caption = 'Spearman correlation at phylum level') %>% 
  kable_styling(full_width = T, position = 'left')

>>>>

Two <- two %>% 
  pluck('shotgun') %>% 
  rename(shotgun_relab = relab) %>% 
  select(-Group) %>% 
    full_join(two %>% 
    pluck('s16S') %>% 
    rename(s16S_relab = relab) %>% 
    select(-Group)) %>%  
  # to uniform the two data
  mutate(s16S_relab = s16S_relab*100) %>% 
  # log10 transform the data add pseudocounto f 0.000001
  mutate(shotgun_relab = log10(shotgun_relab + 0.0001),
         s16S_relab = log10(s16S_relab + 0.0001))

@@

```{r}
Two %>% 
  ggplot(aes(x = shotgun_relab, y = s16S_relab, col = phylum)) +
  geom_point() +
  facet_wrap(~sampleid) +
  geom_abline(intercept = 0, slope = 1, col = 'red', linetype = 'dashed') +
  geom_vline(xintercept = 0, col= 'forestgreen', linetype = 'dashed') +
  geom_hline(yintercept = 0, col= 'forestgreen', linetype = 'dashed') +
  scale_color_manual(values = brewer.pal(n = length(ol_phylum), name = "Dark2")) +
  labs(x = 'log10(Shotgun data)',
       y = 'log10(16S data)',
       title = 'log10 transformed relative abundance comparison between shotgun and 16S data\nat phylum level') +
  ggsave('../figs/correlation/cor_between_16_and_shotgun.jpg', width = 10, height = 10, dpi = 300)

@@

spearman_cor %>% 
  kable(caption = 'Spearman correlation at phylum level') %>% 
  kable_styling(full_width = F, position = 'left')

### 2019-11-08 scripts/21_correlation_genus_bw_16S_shotgun.Rmd
log transform the relab

```{r}
Two %>% 
  ggplot(aes(x = shotgun_relab, y = s16S_relab)) +
    geom_point() +
    facet_wrap(~sampleid) +
  labs(x = 'Shotgun data',
       y = '16S data',
       title = 'Relative abundance comparison between shotgun and 16S data\nat genus level') +
  ggsave('../figs/correlation/cor_between_16_and_shotgun_genus.jpg', width = 10, height = 10, dpi = 300)

spearman_cor <- Two %>% 
  split(.$sampleid) %>% 
  map_dfr(~ round(cor.test(.$shotgun_relab, 
                           .$s16S_relab, 
                           method = 'spearman', 
                           exact = F)$estimate, 2)) %>% 
  gather(key = 'sampleid', value = 'rho') %>% 
  arrange(rho)

spearman_cor %>% 
  kable(caption = 'Spearman correlation at genus level') %>% 
  kable_styling(full_width = T, position = 'left')

>>>>

```{r}
Two %>% 
  ggplot(aes(x = shotgun_relab, y = s16S_relab)) +
    geom_point() +
    facet_wrap(~sampleid) +
  geom_abline(intercept = 0, slope = 1, col = 'red', linetype = 'dashed') +
  geom_vline(xintercept = 0, col= 'forestgreen', linetype = 'dashed') +
  geom_hline(yintercept = 0, col= 'forestgreen', linetype = 'dashed') +
  labs(x = 'log10(Shotgun data)',
       y = 'log10(16S data)',
       title = 'log10 transformed relative abundance comparison between shotgun and 16S data\nat genus level') +
  theme_light()+
  ggsave('../figs/correlation/cor_between_16_and_shotgun_genus.jpg', width = 10, height = 10, dpi = 300)

# plot a single sample's plot to show the details
Two %>% 
  filter(sampleid == '275G') %>% 
  ggplot(aes(x = shotgun_relab, y = s16S_relab)) +
    geom_point() +
    geom_abline(intercept = 0, slope = 1, col = 'red', linetype = 'dashed') +
    geom_vline(xintercept = 0, col= 'forestgreen', linetype = 'dashed') +
    geom_hline(yintercept = 0, col= 'forestgreen', linetype = 'dashed') +
    labs(x = 'log10(Shotgun data)',
         y = 'log10(16S data)',
         title = 'log10(relative abundance) comparison (genus): 275G') +
    theme_light()+
    ggsave('../figs/correlation/275G_cor_between_16_and_shotgun_genus.jpg', width = 5, height = 5, dpi = 300)

# spearman_cor
spearman_cor <- Two %>% 
  split(.$sampleid) %>% 
  map_dfr(~ round(cor.test(.$shotgun_relab, 
                           .$s16S_relab, 
                           method = 'spearman', 
                           exact = F)$estimate, 2)) %>% 
  gather(key = 'sampleid', value = 'rho') %>% 
  arrange(rho)

spearman_cor %>% 
  kable(caption = 'Spearman correlation at genus level') %>% 
  kable_styling(full_width = F, position = 'left')
